---
title: "Bioinformatics Resubmission"
author: '1900357'
date: "2023-08-05"
output: html_document
---

```{r}
print("General RStudio Work")
```
#Below, the "sum()" function is used to add all of the whole numbers from 5 to 55 together. The output value is 1530.
```{r}
sum(5:55)
```
#Following this, the "sumfun()" function to calculate the sum between 5 and 10,20 and 100.
```{r}
sum(5:10)
```
```{r}
sum(5:20)
```
```{r}
sum(5:100)
```
#This relates to the first part of the RStudio Assignment work, as part of candidate number's 1900357 resubmission.
```{r}
print("Fibbonaci Series")
```
#The fibonacci sequence is an integer sequence, with the first two terms as 0 and 1. The following terms are found by adding the previous two terms together. E.g. the next term after 0,1 is 1 (0 + 1 = 1), and so on and so forth. First, the fibonacci sequence must be created and calcuated. Here, we are setting parameters for "n" depending on the value of n, and what it the corresponding output should be. The vector function "c()" is also used to demonstrate the first two terms.
```{r}
fibonacci <- function(n) {
  if (n ==1) {
    return(0)
    } else if (n ==2) {
      return(1)
    } else {
      fib_seq <- c(0,1)
      for (i in 3:n) {
        next_num <- fib_seq[i-1] + fib_seq[i-2]
        fib_seq <- c(fib_seq, next_num)
      }
    return(fib_seq)
  }
}
```

#Here, the nth term is set as 12, as we require the first 12 terms of the fibonacci series. 
```{r}
n_entries <- 12
fib_series <- fibonacci(n_entries)
```

#The "print" function is used to visualise the first 12 terms of the fibonacci sequence.
```{r}
print(fib_series)
```

#The mtcars dataset is a built-in dataset that can be used to help practise visualisation of different parameters, using packages such as "ggplot2". 
```{r}
print("mtcars dataset visualisation")
```
#First, we must retrieve the data.
```{r}
data(mtcars)
```

#The "head" function allows us to visualise the first few lines/ rows of data. This dataset provides information about 32 different cars, based on 11 factors such as miles per gallon (mpg), speed (vs), and the number of gears (gear).
```{r}
head(mtcars)
```
#The "ggplot2" package which is required to visualise data in the form of graphs is installed and loaded for downstream uses. 
```{r}
installed.packages("ggplot2")
library(ggplot2)
```
#The "names" function helps to identify the abbreviations for each of the mtcars dataset rows. The abbreviations will be used donwstream for plotting graph variables.
```{r}
names(mtcars)
```
#The "barplot" function is used to compare the miles per gallon vs the gears. The colour function is used, as well as labels for clarity.
```{r}
barplot(mtcars$mpg, mtcars$gear,
     col='steelblue',
     main='First Graph: MPG vs. Gear',
     xlab='mpg',
     ylab='gear')
```
#The "plot" function is used to plot speed vs. distance.
```{r}
plot(vs ~ disp, data = mtcars, col=2)
```
#The function below is used to create a linear regression model (lm) on the above data. A linear regression model identifies a relationship between the two variables by creating a straight line betwen the two datas.
```{r}
fit <- lm(vs ~ disp, data = mtcars)
```
#The output summary provides a variety of information about the two variables, including the p-values. The p-values (right column under coefficients and last line) will determine whether the linear regression model is statistically significant. This is also denoted by the number of * by the value. The more '*', the more statistically significant the variable. In this case, both variables have p-values < 0.05, with high statistical significance, and so the linear regression is also significant.
```{r}
summary(fit)
```
#Finally, the "abline" function is used to add the regression line to the plot. 
```{r}
abline(plot(fit),col=4,lwd=6)
```

```{r}
print("mtcars dataset: GGplots and Linear Models 2")
```
#The "names" function was used to remind myself of the different variables available.
```{r}
names(mtcars)
```
```{r}
scatter.smooth(x=cars$speed, y=cars$qsec, main="Breaking ~ Speed")  
```

#The "lm" function is used to create a third graph of speed vs. breaking distance, using the mtcars dataset.
```{r}
third_graph <- lm(qsec ~ vs, data = mtcars)
summary(third_graph)
```

#The output produces several different types of plots. One of the main ones of interest is the normal Q-Q plot - a quantile-quantile plot, in which two sets of data (e.e. breaking distance and speed) are placed against one another to see the type of distribution they produce (i.e normal distribtion (bell curve)). If two sets of data originate from the same distribution, they should form an approximate straight line as seen below by the white dots. 
```{r}
abline(plot(third_graph),col=3,lwd=2)
```

